# File of AUC calculation for all three compliance probabilities

# Takes 7 minutes to run once for the test set of n = 10,000 participants and K = 6 time points

# File uses true parameter values to calculate AUC, but parameter estimates can be inserted

### Set the number of participants and time points in the test set
```{r}
n = 10000; t = 6
```

### Source the functions file
```{r,results="hide",message=FALSE,warning=FALSE}
source("C:/Users/rlpet/Documents/GitHub/Longitudinal_Noncompliance_Detection/functions.R")
```

### Generate a large test set as Compound Symmetry
```{r}
set.seed(1)
rs = rnorm(n,0,sqrt(gamma_t))
rs2 = rnorm(n,0,sqrt(tau_t))
b_mat = matrix(NA, n, t)
c_mat = matrix(NA, n, t)
for(j in 1:t){
	c_mat[,j] = rbinom(n,1,pnorm(beta0_t+rs))
	b_mat[,j] = rnorm(n,alpha0_t+alpha1_t*c_mat[,j]+rs2,sqrt(sigma_t))
}
```

### To generate data as AR(1), uncomment this code
```{r}
# c_mat = ps[sample(1:2^t,n,replace=TRUE,prob=probs_AR1),]
# b_mat = t(apply(c_mat,1,bgen))
```

### Replicate the biomarker 2^t times for the 2^t compliance patterns
```{r}
bam = b_mat[rep(1:nrow(b_mat),rep(2^t,n)),]
```

### Calculate the marginal compliance probabilities
```{r}
probs = apply(lu_matr,1,pfun,beta0=beta0_t,
		s_mat=matrix(c(1+gamma_t,rep(c(rep(gamma_t,t),1+gamma_t),t-1)),
		nrow=t),time=t)[rsums]
```

### Calculate the densities of the biomarker
```{r}
fs = dmvnorm(bam-alpha0_t-alpha1_t*cam,z, 
	matrix(c(tau_t+sigma_t,rep(c(rep(tau_t,t),tau_t+sigma_t),t-1)),
	nrow=t))
```

### Calculate the weights
```{r}
m = matrix(probs*fs,nrow=2^t)
wvec = as.vector(t(t(m)/colSums(m)))
```

### Calculate the probabilities of compliance at the last time point
```{r}
ptall = colSums(matrix(wvec[seq(2,2^t*n,2)],nrow=2^(t-1)))
```

### Calculate the probabilities of compliance at all time points
```{r}
pall = wvec[seq(2^t,2^t*n,2^t)]
```

### Calculate the AUC for the probability of compliance at the last time point
```{r,message=FALSE,warning=FALSE}
auc(c_mat[,t],ptall)[1]
```

### Compare with Boatman et al.
```{r,message=FALSE,warning=FALSE}
auc(c_mat[,t],b_mat[,t])[1]
```

### Calculate the AUC for the probability of compliance at all time points
```{r,message=FALSE,warning=FALSE}
auc(rowProds(c_mat),pall)[1]
```

### Compare with Boatman et al.
```{r,message=FALSE,warning=FALSE}
p_mat = matrix(NA,nrow=n,ncol=t)
for(i in 1:t){
	bi = rep(b_mat[,i],each=2)
	probsb = c(pnorm(beta0_t,lower.tail=FALSE),
		pnorm(beta0_t))
	fsb = dnorm(bi-alpha0_t-alpha1_t*rep(c(0,1),n),sd=tau_t+sigma_t)
	mb = matrix(probsb*fsb,nrow=2)
	p_mat[,i] = as.vector(t(t(mb)/colSums(mb)))[seq(2,2*n,2)]
}
auc(rowProds(c_mat),rowProds(p_mat))[1]
```

### Define objects needed for prediction
```{r}
camk1 = psk1[rep(1:2^(t-1),n),]
b_matk1 = b_mat[,1:(t-1)]
bamk1 = b_matk1[rep(1:nrow(b_matk1),rep(2^(t-1),n)),]
```

### Calculate the marginal compliance probabilities for the first five time points
```{r}
probsk1 = apply(lu_matr1,1,pfun,beta0=beta0_t,
		s_mat=matrix(c(1+gamma_t,rep(c(rep(gamma_t,t-1),1+gamma_t),t-2)),
		nrow=t-1),time=t-1)[rsums1]
```

### Calculate the densities of the biomarker for the first five time points
```{r}
fsk1 = dmvnorm(bamk1-alpha0_t-alpha1_t*camk1,zk1, 
		matrix(c(tau_t+sigma_t,rep(c(rep(tau_t,t-1),tau_t+sigma_t),t-2)),
		nrow=t-1))
```

### Gather the information needed for compliance prediction for each participant
```{r}
sup_m = lapply(split(cbind(bam[,1:(t-1)],probs,
	rep(colSums(matrix(probsk1*fsk1,nrow=2^(t-1))),each=2^t)),
	rep(1:n,each=2^t)),matrix,ncol=t+1)
```

### Construct the functions needed for compliance prediction for each participant
```{r}
fns = lapply(sup_m,omf,a0=alpha0_t,a1=alpha1_t,ta=tau_t,sa=sigma_t)
```

### Calculate the prediction probabilities of compliance at the future time point
```{r,cache=TRUE}
pk = sapply(fns,int)
```

### Calculate the AUC for the prediction probability of compliance at the future time point
```{r,message=FALSE,warning=FALSE}
auc(c_mat[,t],pk)[1]
```

### Construct the calibration line
```{r}
mk = matrix(c(c_mat[,t],pk),ncol=2)
dk = mk[order(mk[,2]),]
pkd = colMeans(matrix(dk[,2],ncol=10))
ckd = colSums(matrix(dk[,1],ncol=10))/(n/10)
```

### Plot the calibration line with 45 degree line
```{r}
plot(1,type="n",xlab="Predicted Probability",ylab="Observed Probability",ylim=c(0,1),xlim=c(0,1))
abline(0,1,col="red",lwd=3)
lines(pkd,ckd,lwd=3)
```