# File of parameter estimation with a left-censored biomarker

# Takes 10 minutes to run one simulation for n = 1000 participants and K = 3 time points

### Set the number of participants and time points in the data set
```{r}
n = 1000; t = 3
```

### Source the functions file
```{r,results="hide",message=FALSE,warning=FALSE}
source("C:/Users/rlpet/Documents/GitHub/Longitudinal_Noncompliance_Detection/functions.R")
```

### Generate data as Compound Symmetry
```{r}
set.seed(1)
rs = rnorm(n,0,sqrt(gamma_t))
rs2 = rnorm(n,0,sqrt(tau_t))		
b_mat = matrix(NA, n, t)
c_mat = matrix(NA, n, t)
for(j in 1:t){
	c_mat[,j] = rbinom(n,1,pnorm(beta0_t+rs))
	b_mat[,j] = rnorm(n,alpha0_t+alpha1_t*c_mat[,j]+rs2,sqrt(sigma_t))
}
```


### To generate data as AR(1), uncomment this code
```{r}
# c_mat = ps[sample(1:2^t,n,replace=TRUE,prob=probs_AR1),]
# b_mat = t(apply(c_mat,1,bgen))
```

### Left-censor biomarker data that is below the limit of quantification (LOQ)
```{r}
b_mat = ifelse(b_mat<loq,loq,b_mat)
m = t(apply(b_mat,1,sort))		
rows = apply(m,1,f2)
pseudo = cbind(1:nrow(m),rows)
b_mat2 = m[pseudo[order(pseudo[,2]),][,1],]
```

### Replicate the longitudinal biomarker 2^t times for the 2^t compliance patterns
```{r}
bam = b_mat2[rep(1:n,rep(2^t,n)),]
inds_mat = ifelse(bam<=loq,0,1)
```

### Set starting values of EM algorithm at truth
```{r}
oe = c(alpha0_t,alpha1_t,tau_t,sigma_t,beta0_t,gamma_t)
rel_diff = 10; iter = 0; oinc = 1e6; s = -1
```

### EM algorithm assuming Compound Symmetry covariance structures
```{r,cache=TRUE}
while(rel_diff > 1e-05){
	a0 = oe[1]; a1 = oe[2]; tau = oe[3]; sig = oe[4]
	b0 = oe[5]; gam = oe[6]
	
	# E-step: Calculate the marginal compliance probabilities
	probs = apply(lu_matr,1,pfun,beta0=b0,
		s_mat=matrix(c(1+gam,rep(c(rep(gam,t),1+gam),t-1)),
		nrow=t),time=t)[rsums]

	# E-step: Calculate the densities of the biomarker with left-censoring
	fs = bno(c(a0,a1,log(tau),log(sig)),bam,lg=FALSE,ft=c,tot=c,neg=1,
		time=t,kset=1:(t-1))

	# E-step: Calculate the weights
	m = matrix(probs*fs,nrow=2^t)
	wvec = as.vector(t(t(m)/colSums(m)))
	
	# M-step: Fit the biomarker mixed effects model
	init = c(a0,a1,log(tau),log(sig))
	e1 = optim(par=init,fn=bno,bset=bam,wv=wvec,
		time=t,kset=1:(t-1),method="BFGS")$par
			
	# M-step: Fit the compliance mixed effects model
	init2 = c(b0,log(gam))
	e2 = optim(par=init2,fn=pnlog_lik,w=wvec,method="BFGS")$par
	
	# Calculate the relative difference convergence criterion
	rel_diff = sum(abs((c(e1[1:2],exp(e1[3:4]),e2[1],exp(e2[2]))-oe)/oe))
	
	# Update the number of iterations and parameter estimates
	iter = iter + 1
	oe = c(e1[1:2],exp(e1[3:4]),e2[1],exp(e2[2]))
	a0 = oe[1]; a1 = oe[2]; tau = oe[3]; sig = oe[4]
	b0 = oe[5]; gam = oe[6]
	
	# Calculate the negative incomplete log likelihood
	# The negative log likelihood should decrease with each iteration
	probsn = apply(lu_matr,1,pfun,beta0=b0,
		s_mat=matrix(c(1+gam,rep(c(rep(gam,t),1+gam),t-1)),
		nrow=t),time=t)[rsums]
	fsn = bno(c(a0,a1,log(tau),log(sig)),bam,lg=FALSE,ft=c,tot=c,neg=1,
		time=t,kset=1:(t-1))
	ninc = -sum(log(colSums(matrix(probsn*fsn,nrow=2^t))))

	# Compare the new and previous negative incomplete log likelihood
	s = sign(ninc-oinc)
	
	# Update the negative incomplete log likelihood
	oinc = ninc
}
```

### Check that negative incomplete log likelihood decreased with each iteration
```{r}
s
```

### Compare parameter estimates with truth
```{r}
res = rbind(c(beta0_t,sqrt(gamma_t),alpha0_t,alpha1_t,sqrt(tau_t),sqrt(sigma_t)),
	c(oe[5],sqrt(oe[6]),oe[1:2],sqrt(oe[3:4])))
colnames(res) = c("beta0","gamma","alpha0","alpha1","tau","sigma")
rownames(res) = c("truth","estimate")
print(res)
```